---
title: "20190805_PlotIntronPositions"
author: "Ben Fair"
date: "8/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

```{r}
library(tidyverse)
```


```{r}

files <- list.files(path="../output/SJoutAnnotatedAndIntersected", pattern="*.tab.gz", full.names=TRUE, recursive=FALSE)
SampleName<-gsub("../output/SJoutAnnotatedAndIntersected/(.+).tab.gz", "\\1", files, perl=T)
SampleName

MergedData <- data.frame()
for (i in seq_along(files)){
    CurrentDataset <- read.table(files[i], sep='\t', col.names=c("chr", "start", "stop", "name", "score", "strand", "ASType", "geneChr", "geneStart", "geneStop", "gene", "geneScore", "geneStrand", "Overlap")) %>%
      select(ASType, start, stop, geneChr, geneStart, geneStop, score, strand) %>%
      mutate(samplename=SampleName[i]) %>%
      filter(geneChr != ".") %>%
      mutate(Rel5SplicePos = case_when(
        strand=="+" ~ ((start-geneStart)/(geneStop-geneStart)),
        strand=="-" ~ ((geneStop-stop)/(geneStop-geneStart))
      )) %>%
      mutate(Rel3SplicePos = case_when(
        strand=="+" ~ ((stop-geneStart)/(geneStop-geneStart)),
        strand=="-" ~ ((geneStop-start)/(geneStop-geneStart))
      )) %>%
      filter(Rel5SplicePos<=1 & Rel5SplicePos>=0) %>%
      filter(Rel3SplicePos<=1 & Rel3SplicePos>=0)
    MergedData<-rbind(MergedData,CurrentDataset)
}

SamplesToPlot <- c("Sultan_polyA_Total", "Sultan_rRNADeplete_Total", "Sultan_rRNADepelete_nuclear", "Sultan_rRNADepelete_cytoplasmic", "18862_cheRNA_1","NA18862_argonne", "19201_cheRNA_1", "18913_cheRNA_2")

MergedData %>%
  group_by(samplename, ASType) %>%
  summarise(a_sum=sum(score)) %>%
  group_by(samplename) %>%
  mutate(FractionSpliceType = a_sum/sum(a_sum)*100) %>%
  filter(ASType!="AnnotatedSpliceSite") %>%
  ggplot(aes(x = samplename, y = FractionSpliceType, fill = ASType)) +
  geom_col() +
  ylab("Percent unannotated splicing") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


#Where are called splice sites
MergedData %>%
  filter(samplename %in% SamplesToPlot) %>%
  filter(ASType!="AnnotatedSpliceSite") %>%
  filter(score>0) %>%
ggplot(aes(x=Rel3SplicePos, color=samplename)) +
  geom_density(adjust=2, size=1) +
  xlab("Relative position of unannotated splice site") +
  theme_bw()

#Same, but weighted by RPM for junction, filtering out extreme outliers (>10 counts)
MergedData %>%
  filter(samplename %in% SamplesToPlot) %>%
  filter(ASType!="AnnotatedSpliceSite") %>%
  filter(score>0) %>%
  filter(score<10) %>%
  group_by(samplename) %>%
  mutate(a_sum=sum(score)) %>%
  ungroup() %>%
ggplot(aes(x=Rel5SplicePos, color=samplename)) +
  geom_density(adjust=2, aes(weight=score/a_sum)) +
  theme_bw()

# MergedData %>%
#   filter(samplename %in% SamplesToPlot) %>%
#   filter(ASType!="AnnotatedSpliceSite") %>%
#   filter(score>0) %>%
#   group_by(samplename) %>%
#   mutate(a_sum=sum(score)) %>%
#   mutate(frac=score/a_sum)%>%
#   summarise(max=max(frac))


```

Ok, so nuclear (and chromatin associated) fractions have more unannotated splicing by a factor of about 1.5X to 2X, and of those unannotated splice sites seem very slightly biased towards the 5' end of genes.

As a control, I should make the same metaplots for annotated splice sites

```{r}
#Where are splice events
MergedData %>%
  filter(samplename %in% SamplesToPlot) %>%
  filter(ASType=="AnnotatedSpliceSite") %>%
  filter(score>0) %>%
  ggplot(aes(x=Rel3SplicePos, color=samplename)) +
    geom_density(adjust=2, size=1) +
    xlab("Relative position of annotated splice site") +
    theme_bw()

#Same but weighted by junction RPM
ToPlot <- MergedData %>%
  filter(samplename %in% SamplesToPlot) %>%
  filter(ASType=="AnnotatedSpliceSite") %>%
  filter(score>0) %>%
  filter(score<500) %>%
  group_by(samplename) %>%
  mutate(a_sum=sum(score)) %>%
  ungroup()
ggplot(ToPlot, aes(x=Rel3SplicePos, color=samplename)) +
  geom_density(adjust=2, size=1, aes(weight=score/a_sum)) +
  theme_bw()

MergedData %>%
  filter(samplename %in% SamplesToPlot) %>%
  filter(ASType=="AnnotatedSpliceSite") %>%
  filter(score>0) %>%
ggplot(aes(x=log10(score), color=samplename)) +
  geom_density() +
  theme_bw()

```

Ok, good. So the 5' enrichment for splice sites in chromatin-associated RNAs is true for unannotated splice sites but not for annotated splice sites. 
